---
title: "week 2 assignment"
author: "LevanWR"
date: "9/26/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(dplyr)
library(lattice)

```
we will use following libraries:lubridate, dplyr and lattice.


Code for reading in the dataset and/or processing the data:
```{r}
data<-read.csv("../repro/week2/activity.csv")

#add weekdays to data
data$date<-ymd(data$date)
data$weekdays<-weekdays(data$date)

#clear the data from missing values and add weekdays
data_clear<-na.omit(data)
data_clear$date<-ymd(data_clear$date)
data_clear$weekdays<-weekdays(data_clear$date)


#steps each day 
eachday_steps<-data.frame(tapply(data$steps, factor(data$date), sum, na.rm=TRUE))

#average steps daily
eachday_steps_average<-data.frame(tapply(data$steps, factor(data$interval ), mean, na.rm=TRUE))

#separate weekend and weekdays
data_clear$weekPartition<-data_clear$weekdays
t<-which(data_clear$weekdays=="Saturday"|data_clear$weekdays=="Sunday")
tt<-which(data_clear$weekdays!="Saturday"&data_clear$weekdays!="Sunday")
data_clear$weekPartition[t]<-"weekend"
data_clear$weekPartition[tt]<-"weekdays"


```
Histogram of the total number of steps taken each day

```{r}
hist(eachday_steps$tapply.data.steps..factor.data.date...sum..na.rm...TRUE., xlab = "number of steps", main="Histogram of the total number of steps taken each day")
```
 
 
 Mean and median number of steps taken each day
```{r }
mn<-mean(eachday_steps$tapply.data.steps..factor.data.date...sum..na.rm...TRUE., na.rm = TRUE )
md<-median(eachday_steps$tapply.data.steps..factor.data.date...sum..na.rm...TRUE., na.rm = TRUE)
print(mn)
print(md)
```


Time series plot of the average number of steps taken
```{r}
eachday_steps_average$interval<-unique(data$interval)
plot(eachday_steps_average$interval,eachday_steps_average$tapply.data.steps..factor.data.interval...mean..na.rm...TRUE. , ylab = "averge steps ", main = "Time series plot of the average number of steps taken", type = "l", xlab = "days")

```

The 5-minute interval that, on average, contains the maximum number of steps
```{r}
maximum<-max(eachday_steps_average$tapply.data.steps..factor.data.interval...mean..na.rm...TRUE.)
adress<-which(eachday_steps_average==maximum)
eachday_steps_average$interval[adress]
```

#Code to describe and show a strategy for imputing missing data


##there are missing values in steps, so if we calculate length of steps variable with missing values and without it, difference will give us number of that missing values.  

```{r}
length(data$steps)-length(na.omit(data$steps))

```

Histogram of the total number of steps taken each day after missing values are imputed
```{r}

steps_eachday_clear<-tapply(data_clear$steps, factor(data_clear$date), sum)
hist(steps_eachday_clear, main="Histogram of the total number of steps taken each day after missing values are imputed", xlab = "steps each day total")

```
Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
xyplot(data = data_clear, steps~interval|weekPartition)

```
















